---
title: "Ex 1"
author: "Heidi Al Wakeel"
date: "2023-03-14"
output: html_document
---



```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
# Libraries
library(tidyverse)
library(igraph)
library(janitor)
library(stringi)
```

# Check the data

```{r}
df <-  read.csv("C:/Users/Heidi Al Wakeel/Downloads/Basic_LinkedInDataExport_03-09-2023/Connections.csv")
df <- df %>% 
  select(-c("Email.Address"))
```

# Standardize the name of the companies

There are still some duplicates with inc. or canada. A more systematic way by using the companies name resemblance could be used but that's not the goal of this project and most of the duplicates were fixed by doing these simple fix.

```{r}
# Lower case company name
df <- 
  df %>% 
  mutate(company = tolower(Company))  # lower case
# If no company, write "unknown"
df <- 
  df %>% 
  mutate(Company = replace_na(Company, "None")) %>% 
  mutate(Company = replace(Company, Company=="", "unknown"))
# Remove accents in the column
df$Company <- stri_trans_general(str=df$Company, id="Latin-ASCII")
# Replace everything starting with McGill by just McGill
df <- df %>% 
  mutate(company = replace(Company, str_detect(Company, "mcgill"), "mcgill"))
# Remove "-" and replace with space
df <- df %>% 
  mutate(Company = str_replace(Company, "-", " "))
# PRIVACY
#df %>% head(10)
```

# Get the count of contacts by company

```{r}
count <- df %>% 
  group_by(Company) %>% 
  count() %>% 
  arrange(desc(n))
  
count %>% arrange(desc(n)) %>% head(10) %>% 
  ggplot(aes(y = reorder(Company,n), x=n))+
  geom_col(fill="steelblue") +
  labs(
    x = "Number of contacts",
    y = "Companies",
    title = "Top 10 companies of my connections"
  )
```



```{r}
count
```

# Get the total count

```{r}
total_count = sum(count$n)
print(c("Total connections = ", total_count))
```

# Create the graph

## Create a column with the first and last name

```{r}
df <- df %>% 
  unite(name, c("First.Name", "Last.Name"))
```

## Remove the unknown company contacts from the network
```{r}
df <- df %>% filter(Company!="unknown")
```


## Create the nodes
```{r}
nodes <- df %>% select(c("name", "Company"))
nodes <- nodes %>% rowid_to_column("id")
```

## Create the edges
Left join the id of the contact's name with the same company name

```{r}
edges <- df %>% select(c(name, Company)) %>% 
  left_join(nodes %>% select(c(id,name)), by = c("name"="name"))
edges <- edges %>% left_join(edges, by = "Company", keep=FALSE) %>% 
  select(c("id.x", "id.y", "Company")) %>% 
  filter(id.x!=id.y) # remove the connections between itself
colnames(edges) <- c("x", "y", "Company")
edges %>% head(10)
```

## Create the graph
```{r}
library(tidygraph)
library(ggraph)
graph <- tbl_graph(edges = edges, nodes=nodes, directed = FALSE)
```


## Plot the resulting full graph
```{r}
ggraph(graph, layout = "graphopt") + 
  geom_edge_link(aes(color = Company), show.legend = FALSE) + 
  geom_node_point()+
  theme_graph()
```



## Now, I wanted to take a deeper dive:

```{r }
connections <- read.csv("C:/Users/Heidi Al Wakeel/Downloads/Basic_LinkedInDataExport_03-09-2023/Connections.csv")


connections <- na.omit(connections)


attach(connections)


# Create a table with 
connections$name =  paste(connections$First.Name, substr(connections$Last.Name, start = 1, stop = 1), sep = " ")

connections = connections[, c("name", "Company","Position", "Connected.On")]

# create a frequency table
freq_table = table(connections$Company)
freq_table = sort(freq_table, decreasing = TRUE)
first10= head(freq_table, n = 10)


# Display the list as a table
knitr::kable(first10, col.names = c("Company", "Connections"))
```


### Top 10 Companies and their frequencies
```{r }

# create a bar chart of the frequency table
barplot(first10, main = "Top 10 Companies", 
         ylab = "Number of Connections",
        col = "Pink", las = 2, cex.names = 0.8)

```

### Creating nodes

```{r }

library(tidyverse)

people <- connections %>%
  distinct(name) %>%
  rename(label = name)

companies <- connections %>%
  distinct(Company) %>%
  rename(label = Company)

nodes <- full_join(people, companies, by = "label")
nodes <- rowid_to_column(nodes, "id")
head(nodes)
```


### Creating edges
```{r }

#### Creating edges

edges <- connections[, c("name", "Company")]

edges <- edges %>% 
  left_join(nodes, by = c("name" = "label")) %>% 
  rename(from = id)

edges <- edges %>% 
  left_join(nodes, by = c("Company" = "label")) %>% 
  rename(to = id)

edges <- unique(select(edges, from, to))
head(edges)
```


### Graph using the network library

```{r }

## Building network
library(network)

routes_network <- network(edges,
                          vertex.attr = nodes,
                          matrix.type = "edgelist",
                          ignore.eval = FALSE)
plot(routes_network, vertex.cex = 1,vertex.col="green")
```


### Graph using igraph

```{r }

## igraph

library(igraph)
routes <- graph_from_data_frame(d = edges,
                                       vertices = nodes,
                                       directed = TRUE)

plot(routes,
     vertex.size = 3,
     vertex.label.cex = 0.2,
     edge.arrow.size = 0.01, vertex.col="green")
```






```{r }
library(dplyr)

connections_filtered <- connections %>%
  group_by(Company) %>%
  filter(n() > 1) %>%
  ungroup()

```


### Recreating nodes for companies with more than 1 connection

```{r }

people <- connections_filtered %>%
  distinct(name) %>%
  rename(label = name)

companies <- connections_filtered %>%
  distinct(Company) %>%
  rename(label = Company)

nodes <- full_join(people, companies, by = "label")
nodes <- rowid_to_column(nodes, "id")
```



### Recreating edges

```{r}
#### Creating edges

edges <- connections_filtered[, c("name", "Company")]

edges <- edges %>% 
  left_join(nodes, by = c("name" = "label")) %>% 
  rename(from = id)

edges <- edges %>% 
  left_join(nodes, by = c("Company" = "label")) %>% 
  rename(to = id)

edges <- unique(select(edges, from, to))
```

## Graph using the network library


```{r}
## Building network
library(network)

routes_network <- network(edges,
                          vertex.attr = nodes,
                          matrix.type = "edgelist",
                          ignore.eval = FALSE)
plot(routes_network, vertex.cex = 1, vertex.col="green")
```


### Graph using igraph

```{r}

## igraph

library(igraph)
routes <- graph_from_data_frame(d = edges,
                                       vertices = nodes,
                                       directed = TRUE)

plot(routes,
     vertex.size = 1,
     vertex.label.cex = 0.25,
     edge.arrow.size = 0.05
     )
```











## Final graph with no companies


```{r }
# Filter connections to only include companies with 2 or more employees

contact_count <- connections %>% 
  group_by(Company) %>% 
  summarize(count = n())

Connections <- connections %>%
  inner_join(contact_count, by = "Company") %>%
  filter(count >= 2) %>%
  select(name, Company)

# Create nodes dataframe using tidygraph
nodes <- Connections %>%
  mutate(label = name) %>%
  distinct(label) %>%
  as_tibble() %>%
  select(label)

# Create edges dataframe using tidygraph
edges <- Connections %>%
  left_join(connections, by = "Company") %>%
  filter(name.x != name.y) %>%
  mutate(from = name.x,
         to = name.y) %>%
  select(from, to)

# Create graph using igraph
graph <- graph_from_data_frame(edges, vertices = nodes, directed = FALSE)


```






```{r }
par(mar = rep(1, 4))
options(repr.plot.width = 10, repr.plot.height = 10)
plot(graph, vertex.size = 7, vertex.color = "green", vertex.label.cex = 0.6, edge.color = "gray", edge.width = 2, edge.length=30, vertex.dist = 50)
```

